Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-06-25 paper/paper.tex
revise section 6.3

In $k$-CFA algorithm or the abstract abstract machine shown in Section~\ref{aam},
the call sites @(id 2)@ and @(id 1)@ share the same return flow,
so the invocation of @(id 2)@ returns to both call-sites @[x (id 1)]@ and @[y (id 2)]@.
Therefore, the returned value @2@ for variable @y@ is also propagated to
variable @x@, causing imprecise analysis results to arise.
This return-flow merging is inevitable, even when increasing the context-sensitivity.
If we use the monovariant analysis (i.e., $0$-CFA), the analysis result would be such 
that @x@ and @y@ point to value set @{1, 2}@, because the algorithm does not distinguish 
that @z@ comes from a different call site.
Under $1$-CFA, the algorithm is able to distinguish that variable @z@ of function
@id@ has two different values at two call sites, so variable @y@
would not be polluted by @1@.
However, variable @x@ still points to value set @{1, 2}@, because two call-sites still 
share the same continuation and $1$-CFA is still unable to separate the return-flows. \\

\textit{Call/Return Matching through Refunctionalization and Caching.}
The refunctionalized AAM we obtained has the perfect match on return-flows even though 
we do not have a stack model. The higher-order functions representing continuations 
already connect all the execution in order. In fact, the continuations (call stack) 
of the analyzed language is blended into the call stack of our defining language 
(Scala) through refunctionalization, then the call/return of the analyzed language 
is naturally matched.

In fact, what we started is an AAM with an unbounded stack which already precisely 
matches the calls and returns.
Refunctionalization makes the interpreter be a ``big-step'' semantic artifact, 
then the consequence is we have no places to save the context information. 
But refunctionalization not only forces us to sequentialize the order of nondeterministic 
evaluation through higher-order functions representing continuations, but also drives us
to use a new caching algorithm.

Indeed, the caching algorithm plays an important role to make the analysis computable.

An interesting implication of this is if we apply the caching algorithm with some 
necessary changes to small-step abstract machines with an unbounded stack, 
we are also able to establish a computable and precise call/return match.

>>>>

In $k$-CFA algorithm or the abstract abstract machine shown in Section~\ref{aam},
the call sites @(id 2)@ and @(id 1)@ share the same return flow, so the invocation
of @(id 2)@ returns to both call-sites @[x (id 1)]@ and @[y (id 2)]@.  Therefore,
the returned value @2@ for variable @y@ is also propagated to variable @x@, causing
imprecise analysis results to arise. This return-flow merging is inevitable, even when
increasing the context-sensitivity. Even worse, if we use the monovariant analysis (i.e.,
$0$-CFA), the analysis result would be such that @x@ and @y@ point to value set @{1, 2}@,
because the algorithm does not distinguish that @z@ (the argument of @id@) comes from a
different call site. Under $1$-CFA, the algorithm is able to distinguish that variable
@z@ has two different values at two call sites, so variable @y@ would not be polluted by
@1@. However, variable @x@ still points to value set @{1, 2}@, because two call-sites still
share the same continuation and $1$-CFA is still unable to separate the return-flows. \\

\textit{Call/Return Matching through Refunctionalization and Caching.}
The refunctionalized AAM we obtained has the perfect match on return-flows even though
we do not have a stack model. The higher-order functions representing continuations
already connect all the executions in order. In fact, the continuations (call stack)
of the analyzed language is blended into the call stack of our defining language
(Scala) through refunctionalization, then the call/return of the analyzed language
is naturally matched.

In fact, what we started is an AAM with an unbounded stack which already precisely
matches the calls and returns. Refunctionalization makes the interpreter be a ``big-step''
semantic artifact, then the consequence is we have no places to explicitly save the context
information. But refunctionalization not only forces us to sequentialize the order of
nondeterministic evaluation through higher-order functions representing continuations,
but also drives us to use a new caching algorithm.

Indeed, the caching algorithm plays an important role to guarantee the analysis will
terminate eventually.
An interesting implication of this is if we apply the caching algorithm with some 
necessary changes to small-step abstract machines with an unbounded stack, we are 
also able to establish a computable and precise call/return match.
